Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2015-04-05 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Continuing with Ch4

According to the United States Geological Survey, most gypsum in the United States is used to make wallboards for homes and other structures.  A typical home in the United States contains more than 7 metric tons of gypsum ref USGS website http://minerals.usgs.gov/minerals/pubs/commodity/gypsum/.  As defined by ASTM, gypsum wallboard is gypsum board designed for use on walls, ceilings, or partitions and that affords a surface suitable to receive decoration  ref ATSM 1396. The ASTM standard further describes gypsum wallboard as having a noncombustible core, surfaced with paper bonded to the core. By weight, wallboard is composed of more than 85~\% gypsum, less than 10~\% cellulose, less than 3~\% starch and approximately 1~\% of both crystalline silica and fibrous glass.  Chemically gypsum can be described as calcium sulfate dihydrate, (CaSO$_{4}$2H$_{2}$O) .  By weight pure gypsum is approximately 79~\% calcium sulfate and 21~\% chemically bound water ref Lawson 1977.

The gypsum wallboard used in these experiments is referred to as ``regular'' as opposed to a gypsum wallboard `Type C' or `Type X' which are used in fire rated assemblies.   The gypsum wallboard used in these experiments was purchased from same manufacturer in an effort to have the most similar panels for use between the experiments. Each 1.22~m by 2.44~m by 12.7~mm thick panel weighed approximately 30.5~kg.  The physical features of the gypsum wallboard panels are given in the table below.

@@

After receipt of the gypsum wallboard panels, the panels were laid out on the floor of the NIST Large Fire Research Laboratory, finished paper side up and painted with a roller.  The first coat was a latex primer, http://buyat.ppg.com/rep_pafpainttools_files/Pghpaints/tdb/78200.pdf.  Once the primer dried a top coat of white latex paint was applied http://buyat.ppg.com/rep_pafpainttools_files/Olympic/tdb/78010.pdf .  After the panels dried, they were weighed. The average mass of paint on the finished surface of the gypsum wallboard was 0.33~kg. Then the paint thickness was estimated an ultra sonic XXXX cite - in office, the thickness of the paint varied from 0.05~mm to 0.13~mm.

Some of the panels were then cut-up into samples think about this perhaps this goes into the FP Comp

@@

The first series of experiments were conducted with the natural gas burner.

>>>>

According to the United States Geological Survey, most gypsum in the United States is used to make wallboards for homes and other structures.  A typical home in the United States contains more than 7 metric tons of gypsum ref USGS website http://minerals.usgs.gov/minerals/pubs/commodity/gypsum/.  As defined by ASTM, gypsum wallboard is gypsum board designed for use on walls, ceilings, or partitions and that affords a surface suitable to receive decoration~\cite{ASTM_C1396}. The ASTM standard further describes gypsum wallboard as having a noncombustible core, surfaced with paper bonded to the core. By weight, wallboard is composed of more than 85~\% gypsum, less than 10~\% cellulose, less than 3~\% starch and approximately 1~\% of both crystalline silica and fibrous glass.  Chemically gypsum can be described as calcium sulfate dihydrate, (CaSO$_{4}$2H$_{2}$O) .  By weight pure gypsum is approximately 79~\% calcium sulfate and 21~\% chemically bound water ref Lawson 1977.

The gypsum wallboard used in these experiments is referred to as `regular' as opposed to gypsum wallboard `Type C' or `Type X' which are used in fire rated assemblies.   The gypsum wallboard used in these experiments was purchased from same manufacturer in an effort to have the most similar panels for use between the experiments. Each 1.22~m by 2.44~m by 12.7~mm thick panel weighed approximately 30.5~kg.  The physical features of the gypsum wallboard panels are given in the table below.

@@

After receipt of the gypsum wallboard panels, the panels were laid out on the floor of the NIST Large Fire Research Laboratory, finished paper side up and painted with a roller.  The first coat was a latex primer, http://buyat.ppg.com/rep_pafpainttools_files/Pghpaints/tdb/78200.pdf.  Once the primer dried a top coat of white latex paint was applied http://buyat.ppg.com/rep_pafpainttools_files/Olympic/tdb/78010.pdf .  After the panels dried, they were weighed. The average mass of paint on the finished surface of the gypsum wallboard was 0.33~kg. Then the paint thickness was estimated an ultra sonic XXXX cite - in office.  Over the course of the experiments, random painted panels where checked with the ultra sonic device. The thickness of the paint varied from 0.05~mm to 0.13~mm.  Some of the painted panels were then cut-up into samples for the bench-scale experiments.

@@

In this chapter, as in the previous chapter the averaged peak values for temperatures and heat fluxes will presented at varying distances from the wall. Figures~\ref{IWGB_NG_Temp_FrontSet},\ref{IWGB_NG_Temp_RearCenter} and \ref{IWGB_NG_HF_RearCenter} show the values for the natural gas burner experiments with the gypsum wallboard.
